Quantifying Roughness in SAR Imagery with the Rényi Entropy

### 2025-03-19 Text/Articles/RENYI_ENTROPY_SAR_2025/R0-Renyi-entropy-sar.qmd
Returned with comments

---
title: Rényi Entropy-Based Heterogeneity Detection in SAR Data
format:
  ieee-pdf:
    pdf-engine: pdflatex # xelatex
    keep-tex: true  
    #conference: true # comment this line to use journal
    #journaltype: conference # comment this line to use journal
    fig-cap-location: bottom # to crossref figure
    link-citations: true
    colorlinks: true
    linkcolor: black # equations
    citecolor: black # cites
    urlcolor: black
  #ieee-html: default

author:
  #- id: 
  - name: Janeth Alpala, Abraão D.\ C.\ Nascimento 
    affiliations:
      - name: Universidade Federal de Pernambuco
        department: Departamento de Estatística
        city: Recife 
        country: PE, Brazil
        postal-code: 50670-901
      #- name: Unknown affiliation
   # orcid: 
    email: janeth.alpala@ufpe.br, abraao@de.ufpe.br
    #url: 
    #membership: "Member, IEEE"
    #attributes:
    #  corresponding: true
    #photo: 
    #bio: |
    #  Use `IEEEbiography`  with figure as  option and
    #  the author name as the argument followed by the biography text.
  - name: Alejandro C.\ Frery
    affiliations:
      - name: Victoria University of Wellington
        department: School of Mathematics and Statistics
        city: Wellington 
        country: New Zealand
        postal-code: 6140
      #- name: Unknown affiliation
    #orcid: 
    email: alejandro.frery@vuw.ac.nz
    # bio: |
    #   Use `IEEEbiographynophoto` and the author name
    #   as the argument followed by the biography text.
    # note: "Template created June 23, 2023; revised `r format(Sys.Date(),format='%B %d, %Y')`."
abstract: |
  Synthetic aperture radar (SAR) systems have already been successfully used to solve remote sensing problems. A disadvantage of SAR images is the presence of speckle, which is fully developed in homogeneous areas and gamma-distributed in these scenes. In heterogeneous areas, the intensity values are $\mathcal{G}^0_I$-distributed. 
  In this way, the identification of roughness SAR areas (as opposed to homogeneous ones) is an important task. 
  In this work, we propose a family of hypothesis tests driven by an order parameter to identify roughness features in SAR intensity data using Rényi entropy. 
  In particular, we use a nonparametric estimator for the Rényi entropy and investigate some of its properties. 
  As a practical evaluation method, we develop $p$-value maps on which one can observe both (i) the heterogeneous evidence change per texture and (ii) the prediction of homogeneous and heterogeneous categories. 
  The results are in favor of the Rényi-based heterogeneity detector compared to the one based on Shannon entropy.
keywords: [Rényi entropy, Gamma distribution, heterogeneity, SAR,  hypothesis tests]

@@

# Introduction
[S]{.IEEEPARstart}[ynthetic]{}
 Aperture Radar (SAR) technology has become essential for a wide range of applications\ [@Yu2023;@Akbarizadeh2012;@Mondini2021]. It provides high-resolution data that is independent of sunlight and operates in a variety of weather conditions to facilitate global Earth monitoring\ [@Zeng2020].
Despite its advantages, the effective use of SAR data depends on understanding its statistical properties because this data is affected by speckle, a noise-like interference effect\ [@Argenti2013; @Choi2019]. In intensity format, speckle is typically non-Gaussian, and its presence complicates subsequent image analysis tasks.

The $\mathcal{G}^0_I$ distribution has been shown to be an effective law for SAR intensity data, as it captures different levels of roughness. A limiting case is the Gamma distribution, which arises when speckle is fully developed, indicating homogeneous regions. 
Although these assumptions offer flexibility in describing SAR intensities, selecting the appropriate distribution can be challenging for two main reasons: the small sample sizes used in practical applications and the inherent difficulties associated with parameter estimation. 
These challenges complicate model selection and highlight the need to explore alternative statistical approaches.

Entropy measures have gained attention as valuable statistical tools for analyzing SAR data, with applications in edge detection\ [@Nascimento2014], segmentation\ [@Nobre2016], classification\ [@Cassetti2022], and noise reduction\ [@Chan2022].
Traditionally, Shannon entropy\ [@Shannon1948] has been widely used to quantify uncertainty and disorder in data. However, this study explores an alternative information measure: Rényi entropy—a generalization of Shannon’s formulation. 
This more flexible approach provides additional insights for identifying heterogeneity, making it a promising tool for enhancing SAR image analysis.

In this work, we propose a statistical test based on a non-parametric estimator of Rényi entropy to identify heterogeneous regions in SAR data.
The test assesses whether the observed Rényi entropy in a given region significantly differs from its expected theoretical value under the assumption of homogeneity.
Compared to our previous approach  using Shannon entropy\ [@Frery2024], the Rényi-based approach improves the detection of heterogeneity in real SAR images.

The remainder of this article is organized as follows.
Section\ \ref{sec:pre} outlines an overview of statistical models for SAR intensity data and introduces Rényi entropy.
Section\ \ref{sec:met} describes the proposed hypothesis test and the formulation of the test statistic. 
Section\ \ref{sec:app} evaluates the performance of the test using real SAR data.
Finally, the concluding remarks are drawn in Section\ \ref{sec:conclusion}.

@@

The main distributions considered for SAR intensity data are the $\Gamma_{\text{SAR}}$ distribution, which is suitable for fully developed speckle, and the $\mathcal{G}^0_I$ distribution, which is able to describe roughness\ [@Frery1997]. These distributions are characterized by the following probability density functions (pdfs):
\begin{equation}
	f_{\Gamma_{\text{SAR}}}\bigl(z;L, \mu \bigr) 
    = \frac{L^L}{\Gamma(L)\,\mu^L} z^{L-1} 
    \exp \biggl(-\frac{Lz}{\mu}\biggr)
    \mathbbm 1_{\mathbbm R_+}(z), \label{E:gamma1}
\end{equation}
and
\begin{multline}
    f_{\mathcal{G}^0_I}\bigl(z; \mu, \alpha, L \bigr) 
    = \frac{L^L\,\Gamma(L-\alpha)}
    {\bigl[-\mu(\alpha+1)\bigr]^{\alpha} \Gamma(-\alpha)\,\Gamma(L)}\\
    \frac{z^{L-1}}
    {\bigl[-\mu(\alpha+1)+Lz\bigr]^{L-\alpha}}
    \mathbbm 1_{\mathbbm R_+}(z), \label{E:gi01}
\end{multline}
where $\mu > 0$ is the mean,
$\alpha < 0$ measures the roughness, $L \geq 1$ is the number of
looks, $\Gamma(\cdot)$ is the gamma function, and
$\mathbbm 1_{A}(z)$ is the indicator function of the set $A$. 
As demonstrated by\ [@Frery1997], the $\Gamma_{\text{SAR}}$  model is a particular case of the $\mathcal{G}^0_I$ distribution. Specifically, for a given $\mu$ fixed,
$$
f_{\mathcal{G}^0_I}\big(z; \mu, \alpha, L\big)
\longrightarrow 
f_{\Gamma_{\text{SAR}}}(z;L, \mu) 
$$
when $\alpha\to-\infty$.

@@

Introduced by Alfréd Rényi in 1961\ [@renyi1961measures], this measure generalizes several well-known entropies, including the Shannon entropy\ [@Ribeiro2021]. For a continuous random variable $Z$ with pdf $f(z)$, the Rényi entropy of order $\lambda$, with $\lambda > 0$ and $\lambda \neq 1$, is defined as:
\begin{equation}
\label{E:entropy2}
H_\lambda(Z) = \frac{1}{1 - \lambda} \ln \int_{-\infty}^{\infty} [f(z)]^\lambda \, dz.
\end{equation}
Using \eqref{E:entropy2}, we derive closed-form expressions for the Rényi entropy of the $\Gamma_{\mathrm{SAR}}$ and $\mathcal{G}^0_I$ distributions:
\begin{multline}
\label{eq-HGammaSAR}
H_\lambda\bigl(\Gamma_{\text{SAR}}(L, \mu)\bigr)
= 
\ln \mu - \ln L + \frac{1}{1-\lambda}
\Bigl[
  -\lambda\,\ln\Gamma(L) \\  + \ln\Gamma\bigl(\lambda(L-1)+1\bigr)  - \bigl(\lambda(L-1)+1\bigr)\,\ln\lambda
\Bigr],
\end{multline}
and
\begin{multline}\label{eq-HGI0}
H_\lambda\bigl(\mathcal{G}^0_I(\mu, \alpha, L)\bigr)=\ln\mu-\ln L +\ln(-1 - \alpha)\\
+ \frac{1}{\,1 - \lambda\,}
\Bigl[
   \lambda\bigl(
      \ln\Gamma(L - \alpha)
      -\ln\Gamma(-\alpha)
      -\ln\Gamma(L)
   \bigr)\\
   +\ln\Gamma\bigl(\lambda(L - 1) + 1\bigr)
   +\ln\Gamma\bigl(\lambda(-\alpha + 1) - 1\bigr)\\
   -\ln\Gamma\bigl(\lambda(L - \alpha)\bigr)
\Bigr].
\end{multline}
These results are proven in the Appendix.

@@

## Nonparametric Estimation of Rényi Entropy

Nonparametric estimation of $H(Z)$ has been widely studied using spacing-based estimators, which rely on differences between order statistics [@vasicek1976test; @Ebrahimi1994; @Wieczorkowski1999; @AlOmari2019]. 
Recently, Al-Labadi et al.\ [@AlLabadi2024] proposed a nonparametric estimator for Rényi entropy following this approach.

@@

Following Vasicek [@vasicek1976test] and Ebrahimi et al. [@Ebrahimi1994] for Shannon entropy estimation, and using the $m$-spacing density estimator, the Rényi entropy can be estimated as:
\begin{align}
\label{eq:est_R}
\widehat{H}_\lambda(\bm{Z}) = \frac{1}{1 - \lambda} \ln \left[\frac{1}{n} \sum_{i=1}^{n} \left( \frac{c_i m / n}{Z_{(i+m)} - Z_{(i-m)}} \right)^{\lambda - 1} \right].
\end{align}
This estimator is asymptotically consistent, i.e., it converges in
probability to the true value when $m,n\rightarrow\infty$ and
$m/n\rightarrow0$. 
We choose to use the heuristic spacing $m=\left[\sqrt{n}+0.5\right]$.

@@

Based on the results, we find that the optimal value is $\lambda = 0.9$, as it minimizes the MSE while maintaining a low bias. This choice is the same for $L > 1$. However, for $L = 1$, the optimal $\lambda$ tends to be higher (e.g., $\lambda = 3$) to achieve good results. @fig-plotf illustrates the case for $L = 5$.
```{r fig-plotf, echo=FALSE, message=FALSE, warning=FALSE, fig.fullwidth = TRUE, out.width="42%",  fig.pos="hbt", fig.cap="Bias and MSE as a function of $\\lambda$, with $n=49$, $L=5$.", fig.width=7, fig.height=4.5}

@@

Following Refs.\ [@Frery2024;@Alpala2024], we refined the non-parametric entropy estimator $\widehat{H}_{\lambda}$ in \eqref{eq:est_R} reducing its bias with bootstrap, obtaining $\widetilde{H}_{\lambda}$:
$$
\widetilde{H}_{\lambda} = 2\widehat{H}_{\lambda}(\bm{Z}) - \frac{1}{B} \sum_{b=1}^{B} \widehat{H}_{\lambda}(\bm{Z}^{(b)}),
$$
where $B$ is the number of bootstrap replications, and $\bm{Z}^{(b)}$ is the $b$-th resampled dataset obtained by drawing $n$ observations with replacement from $\bm{Z}$.

A Monte Carlo study with $1,000$ samples from $\Gamma_{\text{SAR}}$ ($\mu = 1$, $L=5$, $\lambda = 0.9$), using $B = 200$ replications, confirms that the bootstrap-corrected estimator $\widetilde{H}_{\lambda}$  outperforms the original $\widehat{H}_{\lambda}$ by reducing both bias and MSE for small sample sizes, as shown in @fig-Plot_bias_msef3.

@@

When the null hypothesis holds, this test statistic is expected to be close to zero.
@fig-density_entropyR shows the empirical density of $S(\bm{Z}; L)$, based on $10,000$ simulations of the $\Gamma_{\text{SAR}}$ model, with $\lambda=0.9$ and $L \in \{5,18\}$.
```{r Simulated_densityR, echo=FALSE, message=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")

@@

```
Under the asymptotic properties of the entropy estimator\ [@vasicek1976test], for sufficiently large samples, $S(\bm{Z}; L)$ follows an asymptotic normal distribution:
\begin{equation*}
S(\bm{Z}; L) \overset{d}{\longrightarrow} \mathcal{N}(\mu_S,\,\sigma^{2}_S)\,,
\end{equation*}
where $d$ represents convergence in distribution. Here, $\mu_S  = \mathbb{E}[S(\bm{Z}; L)]$ and $\sigma^{2}_S = \text{Var}[S(\bm{Z}; L)]$ are the theoretical mean and variance of $S(\bm{Z}; L)$.

We define the standardized test statistic:
\begin{equation*}
\varepsilon = \frac{S(\bm{Z}; L) - \hat{\mu}_S}{\hat{\sigma}_S},
\end{equation*}
where $\hat{\mu}_S$ and $\hat{\sigma}_S$ are the empirical mean and standard deviation of $S(\bm{Z}; L)$, obtained from Monte Carlo simulations under the null hypothesis. 
By the central limit theorem, $\varepsilon$ asymptotically follows a standard normal distribution:
\begin{equation*}
\varepsilon \overset{d}{\longrightarrow} \mathcal{N}(0,1)\,.
\end{equation*}
@fig-density_entropyR_standardized0 shows smoothed histograms of standardized test statistic.
Thus, the $p$-values are calculated as $2\Phi(-|\varepsilon|)$,
where $\Phi(\cdot)$ is the cumulative distribution function of the standard normal distribution.
```{r Simulated_densityR_standardized0, echo=FALSE, message=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")

@@

```
In general, hypothesis tests aim to achieve a controlled type I error rate (size) and high test power (sensitivity to departures from $\mathcal{H}_0$).
To assess these properties, we performed a Monte Carlo simulation with $1000$ replications at significance levels  $1\%$, $5\%$, and $10\%$, evaluating the test under the null hypothesis ($\Gamma_{\text{SAR}}$ distribution), varying sample size and values of $L$ for $\mu=1$. 
The observed type I error rates aligned with the nominal values, confirming the test validity.

@@

In this section, we evaluate the proposed test statistic on real SAR data to assess its effectiveness in detecting heterogeneity.
The test was applied to SAR intensity images from several regions: urban, agricultural, and water. In particular, we analyzed three images: the first from London, United Kingdom; the second from the surroundings of Munich, Germany; and the third from  Dublin Port, Ireland, as shown in Figs.\ \ref{fig:london}(a), \ref{fig:munich}(a), and \ref{fig:dublin}(a).
Table\ \ref{tab:table_param} provides detailed acquisition parameters for these images.
\renewcommand{\arraystretch}{3}   
```{r parameters_sar, echo=FALSE, message=FALSE, warning=FALSE}

@@

To evaluate the performance of our proposal, we applied the test statistics $S_{\widetilde{H}_{\lambda}}(\bm{Z}; L)$ and  $S_{\widetilde{H}_{\text{AO}}}(\bm{Z}; L)$, based on Shannon entropy as described in\ [@Frery2024], using a local sliding window of size $7\times7$. 
The resulting $p$-value maps are presented in Figs.\ \ref{fig:london}(b), \ref{fig:london}(d), \ref{fig:munich}(b), \ref{fig:munich}(d), \ref{fig:dublin}(b), and \ref{fig:dublin}(d). 
These maps employ a color gradient, where darker regions indicate areas with higher roughness, while lighter regions correspond to smoother, less textured surfaces.

In order to better illustrate the decision-making process of both tests, binary maps at a significance level of $0.05$ are shown in Figs.\ \ref{fig:london}(c), \ref{fig:london}(e), \ref{fig:munich}(c), \ref{fig:munich}(e), \ref{fig:dublin}(c), and \ref{fig:dublin}(e).

In these binary representations, white areas correspond to $p$-values greater than $0.05$, indicating that there is no statistical evidence to reject the null hypothesis (homogeneous regions). 
In contrast, black areas represent regions where the $p$-values are less than $0.05$, indicating statistically significant heterogeneity.

The Rényi entropy-based test demonstrated greater sensitivity in detecting textural variations compared to the Shannon-based approach, due to the flexibility provided by the parameter $\lambda$. 
When $\lambda \to 1$, Rényi entropy converges to Shannon entropy, resulting in similar heterogeneity detection. 
Our analysis shows that the optimal $\lambda$ value depends on the number of looks: for single-look images like London, $\lambda = 3$ provided the best contrast, while for multi-look images like Munich and Dublin, $\lambda = 0.9$ was more effective.

@@

# Conclusion {#sec:conclusion}

This study presented a new statistical approach based on Rényi entropy for detecting heterogeneity in SAR images, distinguishing between fully developed speckle and heterogeneous clutter. 
The statistics of the proposed test was computed nonparametrically and its performance was evaluated by a Monte Carlo study. 
The results showed that the method effectively controls the probability of type I error while maintaining a high detection performance that improves as the sample size increases.

>>>>

---
title: Quantifying Roughness in SAR Imagery with the Rényi Entropy
format:
  ieee-pdf:
    pdf-engine: pdflatex # xelatex
    keep-tex: true  
    #conference: true # comment this line to use journal
    #journaltype: conference # comment this line to use journal
    fig-cap-location: bottom # to crossref figure
    link-citations: true
    colorlinks: true
    linkcolor: black # equations
    citecolor: black # cites
    urlcolor: black
  #ieee-html: default

author:
  #- id: 
  - name: Janeth Alpala, Abraão D.&nbsp;C.&nbsp;Nascimento 
    affiliations:
      - name: Universidade Federal de Pernambuco
        department: Departamento de Estatística
        city: Recife 
        country: PE, Brazil
        postal-code: 50670-901
      #- name: Unknown affiliation
   # orcid: 
    email: janeth.alpala@ufpe.br, abraao@de.ufpe.br
    #url: 
    #membership: "Member, IEEE"
    #attributes:
    #  corresponding: true
    #photo: 
    #bio: |
    #  Use `IEEEbiography`  with figure as  option and
    #  the author name as the argument followed by the biography text.
  - name: Alejandro C.&nbsp;Frery
    affiliations:
      - name: Victoria University of Wellington
        department: School of Mathematics and Statistics
        city: Wellington 
        country: New Zealand
        postal-code: 6140
        orcid: "0000-0002-8002-5341"
        membership: "Fellow, IEEE"
      #- name: Unknown affiliation
    #orcid: 
    email: alejandro.frery@vuw.ac.nz
    # bio: |
    #   Use `IEEEbiographynophoto` and the author name
    #   as the argument followed by the biography text.
    # note: "Template created June 23, 2023; revised `r format(Sys.Date(),format='%B %d, %Y')`."
abstract: |
  Quantifying surface roughness in synthetic aperture radar (SAR) data is critical for accurate geophysical interpretation and remote sensing applications. 
  We propose a test statistic based on a non-parametric estimation of Rényi entropy to characterize surface roughness from SAR intensity data. 
  The statistic is refined using bootstrap to improve its stability, size and power. 
  This approach enhances roughness quantification by capturing scale-dependent variations and addressing data-driven uncertainty.
  Experimental results demonstrate the robustness of the proposed method in distinguishing roughness patterns, offering a statistically rigorous tool for SAR-based terrain analysis.
  <!-- Synthetic aperture radar (SAR) systems have already been successfully used to solve remote sensing problems.  -->
  <!-- A disadvantage of SAR images is the presence of speckle, which is fully developed in homogeneous areas and gamma-distributed in these scenes.  -->
  <!-- In heterogeneous areas, the intensity values are $\mathcal{G}^0_I$-distributed.  -->
  <!-- In this way, the identification of roughness SAR areas (as opposed to homogeneous ones) is an important task.  -->
  <!-- In this work, we propose a family of hypothesis tests driven by an order parameter to identify roughness features in SAR intensity data using Rényi entropy.  -->
  <!-- In particular, we use a non-parametric estimator for the Rényi entropy and investigate some of its properties.  -->
  <!-- As a practical evaluation method, we develop $p$-value maps on which one can observe both (i) the heterogeneous evidence change per texture and (ii) the prediction of homogeneous and heterogeneous categories.  -->
  <!-- The results are in favor of the Rényi-based heterogeneity detector compared to the one based on Shannon entropy. -->
keywords: [Rényi entropy, Gamma distribution, heterogeneity, SAR, hypothesis tests, bootstrap]

@@

# Introduction
[S]{.IEEEPARstart}[ynthetic]{}
 Aperture Radar (SAR) technology has become essential for a wide range of applications&nbsp;[@Yu2023;@Akbarizadeh2012;@Mondini2021]. 
 It provides high-resolution data that is independent of sunlight and operates in a variety of weather conditions that facilitate global Earth monitoring&nbsp;[@Zeng2020].
Despite its advantages, the effective use of SAR data depends on understanding its statistical properties because this data is affected by speckle, a noise-like interference effect&nbsp;[@Argenti2013; @Choi2019]. In intensity format, speckle is typically non-Gaussian, and its presence complicates subsequent image analysis tasks.

The $\mathcal{G}^0_I$ distribution effectively characterizes SAR intensity data because it captures different levels of roughness. 
A limiting case is the Gamma distribution, which arises when speckle is fully developed, indicating textureless regions. 
Although these assumptions offer flexibility in describing SAR intensities, selecting the appropriate distribution can be challenging for two main reasons: the small sample sizes used in practical applications and the inherent difficulties associated with parameter estimation. 
These challenges complicate model selection and highlight the need to explore alternative statistical approaches.

Entropy measures have gained attention as valuable statistical tools for analyzing SAR data, with applications in edge detection&nbsp;[@Nascimento2014], 
segmentation&nbsp;[@Nobre2016], 
classification&nbsp;[@Cassetti2022], 
and noise reduction&nbsp;[@Chan2022].
Traditionally, Shannon entropy&nbsp;[@Shannon1948] has been widely used to quantify data uncertainty and disorder. 
We explore a more general information measure: Rényi entropy—a generalization of Shannon’s formulation. 
This more flexible approach provides additional insights for identifying heterogeneity, making it a promising tool for enhancing SAR image analysis.

We propose a statistical test based on a non-parametric estimator of Rényi entropy to identify heterogeneous regions in SAR data.
The test assesses whether the observed Rényi entropy significantly differs from its expected theoretical value under the assumption of homogeneity.
Compared to our previous approach using Shannon entropy&nbsp;[@Frery2024], the Rényi-based approach improves the detection of heterogeneity.

The remainder of this article is organized as follows.
Section&nbsp;\ref{sec:pre} overviews the statistical models for SAR intensity data and introduces Rényi entropy.
Section&nbsp;\ref{sec:met} describes the proposed hypothesis test and the test statistic.
Section&nbsp;\ref{sec:app} evaluates the performance of the test using SAR data.
Finally, the concluding remarks are drawn in Section&nbsp;\ref{sec:conclusion}.

@@

The main distributions considered for SAR intensity data are the $\Gamma_{\text{SAR}}$ distribution, which is suitable for fully developed speckle, and the $\mathcal{G}^0_I$ distribution, which is able to describe roughness&nbsp;[@Frery1997]. These distributions are characterized by the following probability density functions (pdfs):
\begin{equation}
	f_{\Gamma_{\text{SAR}}}\bigl(z;L, \mu \bigr) 
    = \frac{L^L}{\Gamma(L)\,\mu^L} z^{L-1} 
    \exp \biggl(-\frac{Lz}{\mu}\biggr)
    \mathbbm 1_{\mathbbm R_+}(z), \label{E:gamma1}
\end{equation}
and
\begin{multline}
    f_{\mathcal{G}^0_I}\bigl(z; \mu, \alpha, L \bigr) 
    = \frac{L^L\,\Gamma(L-\alpha)}
    {\bigl[-\mu(\alpha+1)\bigr]^{\alpha} \Gamma(-\alpha)\,\Gamma(L)}\\
    \frac{z^{L-1}}
    {\bigl[-\mu(\alpha+1)+Lz\bigr]^{L-\alpha}}
    \mathbbm 1_{\mathbbm R_+}(z), \label{E:gi01}
\end{multline}
where $\mu > 0$ is the mean,
$\alpha < 0$ measures the roughness, $L \geq 1$ is the number of
looks, $\Gamma(\cdot)$ is the gamma function, and
$\mathbbm 1_{A}(z)$ is the indicator function of the set $A$. 
As demonstrated by&nbsp;[@Frery1997], the $\Gamma_{\text{SAR}}$  model is a particular case of the $\mathcal{G}^0_I$ distribution. Specifically, for a given $\mu$ fixed,
$$
f_{\mathcal{G}^0_I}\big(z; \mu, \alpha, L\big)
\longrightarrow 
f_{\Gamma_{\text{SAR}}}(z;L, \mu) 
$$
when $\alpha\to-\infty$.

@@

Introduced by Alfréd Rényi in 1961&nbsp;[@renyi1961measures], this measure generalizes several well-known entropies, including Shannon's&nbsp;[@Ribeiro2021]. 
For a continuous random variable $Z$ with pdf $f(z)$, the Rényi entropy of order $\lambda \in \mathbbm R_+ \setminus \{1\}$, is defined as:
\begin{equation}
\label{E:entropy2}
H_\lambda(Z) = \frac{1}{1 - \lambda} \ln \int_{-\infty}^{\infty} [f(z)]^\lambda \, dz.
\end{equation}
Using \eqref{E:entropy2}, we derive closed-form expressions for the Rényi entropy of the $\Gamma_{\mathrm{SAR}}$ and $\mathcal{G}^0_I$ distributions:
\begin{multline}
\label{eq-HGammaSAR}
H_\lambda\bigl(\Gamma_{\text{SAR}}(L, \mu)\bigr)
= 
\ln \mu - \ln L + \frac{1}{1-\lambda}
\Bigl[
  -\lambda\,\ln\Gamma(L)  \\ + \ln\Gamma\bigl(\lambda(L-1)+1\bigr)  - \bigl(\lambda(L-1)+1\bigr)\,\ln\lambda
\Bigr],
\end{multline}
and
\begin{multline}\label{eq-HGI0}
H_\lambda\bigl(\mathcal{G}^0_I(\mu, \alpha, L)\bigr)=\ln\mu-\ln L +\ln(-1 - \alpha)\\
+ \frac{1}{1 - \lambda}
\Bigl[
   \lambda\bigl(
      \ln\Gamma(L - \alpha)
      -\ln\Gamma(-\alpha)
      -\ln\Gamma(L)
   \bigr)\\
   +\ln\Gamma\bigl(\lambda(L - 1) + 1\bigr)
   +\ln\Gamma\bigl(\lambda(-\alpha + 1) - 1\bigr)\\
   -\ln\Gamma\bigl(\lambda(L - \alpha)\bigr)
\Bigr].
\end{multline}
These results are proven in the Appendix.
<!-- ACF Have you checked if it is possible to state the difference between \label{eq-HGI0} and \label{eq-HGammaSAR}? At the end of the day, that is what we want to assess. -->

@@

## Non-parametric Estimation of Rényi Entropy

In order to estimate&nbsp;\eqref{E:entropy2} from a random sample, one may estimate the parameters that index the probability density function $f$, and integrate.
This approach heavily relies on the model assumptions and on the absence of outliers.
Alternatively, one may estimate directly $f$.
Such non-parametric estimation of $H(Z)$ has been widely studied using spacing-based estimators, which rely on differences between order statistics [@vasicek1976test; @Ebrahimi1994; @Wieczorkowski1999; @AlOmari2019]. 
Recently, Al-Labadi et al.&nbsp;[@AlLabadi2024] proposed a non-parametric estimator for Rényi entropy following this approach.

@@

Following Vasicek&nbsp;[@vasicek1976test] and Ebrahimi et al.&nbsp;[@Ebrahimi1994] for Shannon entropy estimation, and using the $m$-spacing density estimator, the Rényi entropy can be estimated as:
\begin{align}
\label{eq:est_R}
\widehat{H}_\lambda(\bm{Z}) = \frac{1}{1 - \lambda} \ln \left[\frac{1}{n} \sum_{i=1}^{n} \left( \frac{c_i m / n}{Z_{(i+m)} - Z_{(i-m)}} \right)^{\lambda - 1} \right].
\end{align}
This estimator is asymptotically consistent, i.e., it converges in
probability to the true value when $m,n\rightarrow\infty$ and
$m/n\rightarrow0$. 
We used the heuristic spacing $m=\left[\sqrt{n}+0.5\right]$.

@@

We found that for $L > 1$ the optimal value is $\lambda = 0.9$, as it minimizes the MSE while maintaining a low bias.
However, for $L = 1$, the optimal $\lambda$ tends to be higher and we chose $\lambda = 3$ to achieve good results. 
@fig-plotf illustrates the case for $L = 5$.

```{r fig-plotf, echo=FALSE, message=FALSE, warning=FALSE, fig.fullwidth = TRUE, out.width="42%",  fig.pos="hbt", fig.cap="Bias and MSE as a function of $\\lambda$, with $n=49$, $L=5$.", fig.width=7, fig.height=4.5}

@@

Following Refs.&nbsp;[@Frery2024;@Alpala2024], we refined the non-parametric entropy estimator $\widehat{H}_{\lambda}$ in&nbsp;\eqref{eq:est_R} reducing its bias with bootstrap, obtaining $\widetilde{H}_{\lambda}$:
$$
\widetilde{H}_{\lambda} = 2\widehat{H}_{\lambda}(\bm{Z}) - \frac{1}{B} \sum_{b=1}^{B} \widehat{H}_{\lambda}(\bm{Z}^{(b)}),
$$
where $B$ is the number of bootstrap replications, and $\bm{Z}^{(b)}$ is the $b$-th resampled dataset obtained by drawing $n$ observations with replacement from $\bm{Z}$.

A Monte Carlo study with $1000$ samples from $\Gamma_{\text{SAR}}$, $\mu = 1$ and $L=5$, using $B = 200$ replications, confirms that the bootstrap-corrected estimator $\widetilde{H}_{\lambda}$, with $\lambda = 0.9$, outperforms the original $\widehat{H}_{\lambda}$ by reducing both bias and MSE for small sample sizes, as shown in @fig-Plot_bias_msef3.
<!-- ACF Remove the minor vertical grid (there is nothing between 49 and 81) -->

@@

When the null hypothesis holds, this test statistic is expected to be close to zero.
@fig-density_entropyR shows the empirical density of $S(\bm{Z}; L)$, based on $10^4$ replications of the $\Gamma_{\text{SAR}}$ model, with $\lambda=0.9$ and $L \in \{5,18\}$.
<!-- ACF 10^4 replications of samples of size?   -->

```{r Simulated_densityR, echo=FALSE, message=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")

@@

Vasicek&nbsp;[@vasicek1976test] proved that, for sufficiently large samples, $S(\bm{Z}; L)$ asymptotically follows a normal distribution:
\begin{equation*}
S(\bm{Z}; L) \overset{d}{\longrightarrow} \mathcal{N}(\mu_S,\,\sigma^{2}_S)\,,
\end{equation*}
where $d$ represents convergence in distribution. Here, $\mu_S  = \mathbb{E}[S(\bm{Z}; L)]$ and $\sigma^{2}_S = \text{Var}[S(\bm{Z}; L)]$ are the theoretical mean and variance of $S(\bm{Z}; L)$.

We define the standardized test statistic:
\begin{equation*}
\varepsilon = \frac{S(\bm{Z}; L) - \hat{\mu}_S}{\hat{\sigma}_S},
\end{equation*}
where $\hat{\mu}_S$ and $\hat{\sigma}_S$ are the empirical mean and standard deviation of $S(\bm{Z}; L)$, obtained by Monte Carlo simulations under the null hypothesis. 
By the central limit theorem, $\varepsilon$ asymptotically follows a standard normal distribution:
\begin{equation*}
\varepsilon \overset{d}{\longrightarrow} \mathcal{N}(0,1)\,.
\end{equation*}
Thus, the $p$-values are calculated as $2\Phi(-|\varepsilon|)$,
where $\Phi(\cdot)$ is the cumulative distribution function of the standard normal distribution.
@fig-density_entropyR_standardized0 shows smoothed histograms of standardized test statistics.

```{r Simulated_densityR_standardized0, echo=FALSE, message=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")

@@

```
In general, hypotheses tests aim to control the Type&nbsp;I error rate (size) with high test power (sensitivity to departures from $\mathcal{H}_0$, low Type&nbsp;II error rate).
To assess these properties, we performed a Monte Carlo simulation with $1000$ replications at significance levels  $1\%$, $5\%$, and $10\%$, evaluating the test under the null hypothesis ($\Gamma_{\text{SAR}}$ distribution), varying sample size and values of $L$ for $\mu=1$. 
<!-- ACF I wish you had used the siunitx package for typing percentages, but not if this is incompatible with the tables -->
The observed Type&nbsp;I error rates align with the nominal values, confirming the test validity.

@@

In this section, we evaluate the proposed test statistic on SAR data to assess its effectiveness in detecting heterogeneity.
We applied the test to images covering urban, agricultural, and water regions. 
We analyzed images from London, United Kingdom; from the surroundings of Munich, Germany; and from  Dublin Port, Ireland, as shown in Figs.&nbsp;\ref{fig:london}(a), \ref{fig:munich}(a), and \ref{fig:dublin}(a).
Table&nbsp;\ref{tab:table_param} provides detailed acquisition parameters for these images.

<!-- ACF They all have HH polarization; the column is not informative, and can be deleted to gain space -->
\renewcommand{\arraystretch}{3}   
```{r parameters_sar, echo=FALSE, message=FALSE, warning=FALSE}

@@

We applied the test statistics $S_{\widetilde{H}_{\lambda}}(\bm{Z}; L)$ and  $S_{\widetilde{H}_{\text{AO}}}(\bm{Z}; L)$, based on Shannon entropy as described in&nbsp;[@Frery2024], using a local sliding window of size $7\times7$. 
<!-- ACF What is $S_{\widetilde{H}_{\text{AO}}}(\bm{Z}; L)$? -->
The resulting $p$-value maps are presented in Figs.&nbsp;\ref{fig:london}(b), \ref{fig:london}(d), \ref{fig:munich}(b), \ref{fig:munich}(d), \ref{fig:dublin}(b), and \ref{fig:dublin}(d). 
These maps employ a color gradient, where darker regions indicate areas with higher roughness, while lighter regions correspond to smoother, less textured surfaces.

Figs.&nbsp;\ref{fig:london}(c), \ref{fig:london}(e), \ref{fig:munich}(c), \ref{fig:munich}(e), \ref{fig:dublin}(c), and \ref{fig:dublin}(e) 
are binary maps at significance level of $0.05$ that 
illustrate the decision-making process.
In these binary representations, white areas correspond to $p$-values greater than $0.05$, indicating that there is no statistical evidence to reject the null hypothesis (homogeneous regions). 
In contrast, black areas represent regions where the $p$-values are less than $0.05$, indicating statistically significant heterogeneity.

The Rényi entropy-based test demonstrated greater sensitivity in detecting textural variations compared to the Shannon-based approach, due to the flexibility provided by the parameter $\lambda$. 
<!-- ACF Where is the Shannon-based approach shown? -->
<!-- When $\lambda \to 1$, Rényi entropy converges to Shannon entropy, resulting in similar heterogeneity detection.  -->

@@

# Conclusions {#sec:conclusion}

This study presented a new statistical approach based on Rényi entropy for detecting heterogeneity in SAR images, distinguishing between fully developed speckle and heterogeneous clutter. 
The statistics of the proposed test was computed non-parametrically and its performance was evaluated by a Monte Carlo study. 
The results showed that the method effectively controls the probability of Type&nbsp;I error while maintaining a high detection performance that improves as the sample size increases.
